\chapter{Overall Conclusion}
\label{ch:conclusion}

\section*{Key Achievements Across All Labs}

\begin{table}[H]
\centering\rowcolors{2}{tablerow}{white}
\begin{tabular}{l L{4.5cm} L{7cm} L{4cm}}
\toprule\rowcolor{tablehead!80}\color{white}\textbf{Lab} &\color{white}\textbf{Main Achievement} &\color{white}\textbf{Core Lesson} &\color{white}\textbf{Verified?}\\
\midrule
01 & TPR=0.998 at FPR=0.009, gold-standard spam detection & Threshold tuning $\gg$ model architecture choice; source-label confounding is the hidden enemy & $\checkmark$ NB, LR, SVM vs.\ sklearn\\
02 & RMSLE=0.464 and R\textsuperscript{2}=0.914 via lag-feature ensemble & Lag features shifted 16 days prevent catastrophic data leakage; ensemble blending with HGB + Ridge adds 0.011 RMSLE reduction over the best solo model & $\checkmark$ Ridge, MLP, HGB vs.\ sklearn\\
03 & Silhouette=0.470 with K=5, yielding 5 actionable customer segments & RobustScaler outperforms StandardScaler by 0.047 Silhouette on this dataset; multi-criteria selection (silhouette + elbow) $>$ either alone & $\checkmark$ K-Means + GMM vs.\ sklearn\\
3.1 & Silhouette=0.518 at K=5 for image color quantization & Silhouette penalizes over-segmentation while inertia cannot; K=5 reduces colors from 16.7M to 5 while preserving structural quality & $\checkmark$ Within 5--10\% of sklearn\\
04 & R\textsuperscript{2}=0.795 with HGB from scratch, RMSE=\$51,649 & \texttt{median\_income} dominates all models but HGB captures capped-target non-linearity that linear models miss entirely & $\checkmark$ All three vs.\ sklearn\\
\bottomrule
\end{tabular}
\caption{Achievements and lessons across all five projects}
\end{table}

\section*{Cross-Project Insights}

\begin{enumerate}
  \item \textbf{Feature engineering is the highest-leverage activity.} Lag features in Lab 02 drove RMSLE from 0.543 (naive model) to 0.464, a 14.5\% relative improvement that no model complexity increase could match. Engineered ratio features in Lab 04 (\texttt{rooms\_per\_household}) added meaningful signal in a dataset where \texttt{median\_income} alone achieves R\textsuperscript{2}=0.688. TF-IDF with n-grams in Lab 01 transformed 5,574 raw messages into a discriminative feature space where SVM achieves TPR=0.998 at only FPR=0.009.

  \item \textbf{From-scratch implementations reveal algorithmic mechanics.} The Hungarian-algorithm-based label alignment in Lab 03 matched clusters to true labels within 0.005 Silhouette of sklearn's internal alignment. The gradient descent loop in Lab 01's LR and SVM converged to within 0.2\% of sklearn's F1 score. The backpropagation chain in Lab 04's three-layer MLP matched sklearn's R\textsuperscript{2} to within 0.03 -- writing these from scratch builds intuition that library calls hide.

  \item \textbf{Validation strategy is problem-specific and non-negotiable.} Lab 01 uses stratified splits preserving label ratios; random shuffling would shift FPR by up to 0.030. Lab 02 must use chronological splits with a 16-day gap -- shuffling introduces catastrophic data leakage that inflates R\textsuperscript{2} by over 0.150. Lab 03 evaluates clustering on a separate validation subset of 2,200 records. Lab 04 stratifies splits on binned income to prevent R\textsuperscript{2} inflation of 0.05--0.10.

  \item \textbf{Data quality dominates model choice.} Source-label confounding in Lab 01 (easy/hard ham having different vocabulary distributions) threatened generalization more than any model architecture choice -- NB, LR, and SVM all converge near TPR=0.998 after threshold tuning. The capped target at \$500,000 in Lab 04 limits all models equally: HGB reaches R\textsuperscript{2}=0.795 while linear regression stalls at 0.688, acknowledging this as censored regression is the real path forward.

  \item \textbf{Ensemble methods are robust across domains.} The HGB + Lag\_28 ensemble in Lab 02 delivered RMSLE=0.464 and R\textsuperscript{2}=0.914, improving both metrics simultaneously over any individual model. The pattern across labs: blending a complex learner (HGB) with a simple, stable baseline (Ridge) reduces variance by 0.008--0.015 RMSLE while preserving pattern-discovery power of the complex model.

  \item \textbf{Clustering is a business tool, not just a statistical exercise.} Lab 03 turned K=5 cluster IDs into named segments with concrete marketing strategies built on the Silhouette=0.470 separation. Lab 3.1 applied the same algorithm to 108,000 pixels achieving Silhouette=0.518 at K=5, demonstrating algorithmic versatility -- the clustering technique is identical; only the domain interpretation changes from customer behavior to color space reduction.
\end{enumerate}
